### 2023-04-11 preregistrations/3_study2/preregistration.Rmd
initiate study 2 folders

In line with previous thinking (e.g., Ellis et al., 2020; Frankenhuis et al., 2016; Frankenhuis & Weerth, 2013), we initially expected people from adversity to perform better on cognitive tasks that require a broad attention scope. A broad attention scope might facilitate vigilance and the detection of subtle changes or peripheral information. However, this might come at a cost, especially for tasks that involve interference. To test these ideas, we ran a pilot study with a Cued Attention, Changed Detection, and Flanker Task (interference) and measured adversity exposure. We expected adversity-exposure would improve performance on the Cued-Attention and Change Detection tasks but hurt performance on the Flanker Task (see the preregistration of the first pilot study).

In contrast to expectations, our pilot data did not find evidence for a broad adversity-related attention scope. In fact, our Drift Diffusion analyses showed that participants with more violence exposure were generally slower to process information – as shown by the drift rate parameters of the Cued Attention and Change Detection Task and the perceptual input parameter of the Flanker Task - and orient their attention – As shown by the non-decision time parameter - across all three tasks. However, on the Flanker Task — where we expected lowered performance — our analyses revealed that people exposed to violence were faster to narrow their attention to the task-relevant information. In other words, instead of maintaining a broad scope of attention, people exposed to violence were able to quickly focus their attention, as shown by the attention shrinking parameter.

These findings are interesting for two reasons. First, the Flanker is a widely used task and people exposed to adversity typically show lowered performance. However, using the DDM to break performance into components, our pilot data suggest that lowered performance may not be caused by problems with attention control or inhibition. Instead, people exposed to adversity might simply extract perceptual information from the stimuli more slowly (as indicated by the perceptual input (p) parameter). This effectively lowers the quality of information that people can use to guide their decision, which will be reflected in lower reaction times. Second, these initial findings suggest that performance might be improved through different types of interventions focused on different cognitive mechanisms. Specifically, people with more adversity exposure might benefit from manipulations that increase the visual quality of stimuli.

The goal of the current study is to replicate the attention findings on the Flanker Task, and to better understand the apparent information processing deficits. Participants completed three versions of the Flanker Task: the standard version, a version with enhanced visual information, and a version with degraded visual information.

Figure \@ref(fig:figure1) shows two possible data patterns that we might expect across the different versions at different levels of violence exposure. First, if the findings reflect a general information processing deficit, we would expect performance to be lower for people with more violence exposure across all versions of the task. Alternatively, lower performance on the standard version might reflect an adaptive trade-off towards more robust cognitive functioning, even in the face of noise or perturbations [@giudice_2018]. In that case, we would expect more stable performance for people with more violence exposure. As a result, they might not benefit as much from enhanced visual information, but at the same time might be able to better maintain performance when presented with degraded information.

```{r figure1, fig.width=6, dpi=600, fig.id = "figure1", fig.cap.style = "Image Caption", fig.cap="Overview of the three Flanker conditions and predicted patterns of results under different cognitive mechanisms. A: The different Flanker Task conditions. The standard version is a typical implementation of the Flanker stimuli. The enhanced version improves the quality of information by increasing stimulus size by 12.5% and including padding (5 pixels) between the arrows. The degraded cognition reduces the quality of information by rotating all arrows at a 45° angle. B: Two potential patterns of the effect of condition on people’s perceptual input as a function of violence exposure. Left plot: Under the deficit model, we would expect people with more violence exposure to perform worse across all three conditions. Right plot: If performance is traded off against higher robustness of the cognitive system for people with more violence exposure (Del Giudice & Crespi, 2018), we would expect cognitive performance in this group to be relatively stable across conditions, perhaps even outperforming people with less violence exposure in the degraded condition."}
knitr::include_graphics("Fig1.png", error = F)
```

@@

1. Investigate the robustness of the primary DDM findings in the pilot study by pooling the pilot data and the data of the standard condition in the current study. 
2. Extent the findings of the pilot study by investigating how Flanker performance changes across conditions as a function of violence exposure. We expect that the manipulations primarily have an effect on perceptual input *p*, thereby also affecting overall reaction times.

# Secondary aims  {#secondary}

1. Investigate the extent to which the [factor structure](#efa) of the unpredictability measures corresponds to the one found in the pilot study. Secondly, investigate how different dimensions of unpredictability relate to Flanker performance.
2) Explore the role of state anxiety, hunger, and sleep deprivation as potential moderators of the relationship between adversity and attention performance. Anxiety might enhance attention performance by making participants more vigilant. Conversely, hunger and sleep deprivation could have a negative effect on performance.
3) Explore bivariate correlations between measures of adversity, attention, and measures of temporal orientation (i.e., impulsivity and future orientation).
4) Explore the correlation between current depressive symptoms and retrospective measures of adversity, which might reflect a negativity bias in recalling past events driven by current depressive symptoms [@nivison_2021].

@@

Participants were 550 US-based individuals between the ages of 18 and 30 (*M* = `r #TBD`, *SD* = `r #TBD`). We decided for this age cut-off to reduce the effect of healthy age-related cognitive decline on speeded cognitive tasks, which becomes more pronounced after age 30 [@salthouse_2010]. Recruitment took place through Prolific Academic [www.prolific.co]. The sample was balanced to include roughly 50% males. Participants were eligible for the study if they were from the United States and if they spoke fluent English. Finally, we used the MacArthur's ladder for perceived social-economic status (SES) as used in Prolific's pre-screening battery to ensure that 50% of the sample perceived their current SES to be low (a score of 4 or below).

We conducted a power simulation using the *faux* package in R [@debruine_2021] to determine the optimal number of participants to include (more information including all simulation code on [Github](https://github.com/StefanVermeent/attention_project/tree/preregistration1/preregistrations/1_pilot/scripts)). Assuming an effect size of $\beta$ = 0.1 and a DDM parameter recovery accuracy of *r* = .85, we estimated that we would need 500 participants to achieve at least 80% power. We sampled a total of 550 participants to account for necessary exclusions (based on our experiences in the pilot study).

@@

We applied several exclusion criteria prior to analyzing the data. First, we excluded participants who did not complete the full study and those who did not complete all three Flanker conditions (*N* = `r #TBD `). Second, we analyzed responses to the attention checks and reversed coded items in the questionnaire part of the experiment. We excluded participants if they missed both attention check items or if they had suspicious response patterns (e.g., consistently endorsing high response options even when some items were reverse coded) (*N* = `r #TBD `). Third, we excluded participants whose screen height was < 700 pixels, or whose screen height was bigger than their screen width (suggesting that they did not complete the experiment on a laptop or desktop pc). Finally, we screened participants' accuracy on the Flanker task within each condition. Using a binomial distribution, we calculated that if a participant would be purely guessing on each trial, they would have a probability of 97.5% of obtaining an overall accuracy of 59.4%. Therefore, we excluded participant's data on a particular condition if their accuracy on that condition was below 60%.

In addition, we screened the reaction times on the Flanker Task. For the Drift Diffusion analyses, it is important that each response is generated by a process of active information accumulation (i.e., a diffusion process, as opposed to guessing). To this end, trials with reaction times < 250 ms or > 3500 ms [@ratcliff_2015] were excluded from the analyses (*N* = `r #TBD `). Participants with more than 10 removed trials were excluded from the analyses (N = `r #TBD `). Finally, we logged whether participants exited full-screen mode and/or engaged with other browser tabs (i.e., blur events) at any point during the Flanker Task. We excluded participants for whom blur events occurred while a Flanker block was ongoing (but not while reading instructions or taking breaks in between conditions). Full-screen exits were included in the [multiverse analysis](#multiverse_plan).

@@

Based on participant feedback in the pilot study, we changed (the wording of) a number of questionnaire items in order to make them more inclusive. We added an item asking about the participant's family composition. In addition, all mentions of 'parent(s)' across questionnaires were changed to 'caregiver(s)' (e.g., "At least one of my caregivers had punishments that were unpredictable"). We also added a question asking about the participant's exact family composition.

@@

The Flanker Task is commonly used as a measure of selective attention and response inhibition [@eriksen_1974]. On each trial of a typical Flanker Task, participants are presented with a set of five arrows pointing either left or right. Their job is to indicate the direction of the central arrow while ignoring the flanking arrows to the left and right (see Figure \@ref(fig:figure1)b). On 50% of the trials, the flanking arrows point in the same direction as the central arrow (i.e., congruent trials), and on the other 50% of the trials they point in the opposite direction (i.e., incongruent trials).

The Flanker Task was programmed in JsPsych version 6.3.1 [@deLeeuw_2015]. We programmed three conditions of the Flanker Task (See Figure \@ref(fig:figure1). In the standard condition, the arrows were 40 pixels in size (0.4 inches) and had zero padding between them. In the 'enhanced' condition, the aim was to increase the quality of visual information provided by the arrows. To this end, we increased the arrow size by 12.5% to 45 pixels (0.45 inches), and increased padding between the arrows to 5 pixels. This increased the width of the Flanker display by 50% with respect to the standard display. In the 'degraded' condition, the aim was to decrease the quality of visual information provided by the arrows. The arrow size and padding were the same as in the standard version, but both flanking and target arrows were rotated 45$^\circ$. Therefore, they provided more noisy information about their horizontal direction. We had no strong *a priori* expectations about how performance of adversity-exposed people might be affected by specific types of stimulus degradation (e.g., arrow rotation, decreasing stimulus contrast).

Participants completed each condition of the Flanker Task separately in different blocks. The presentation order of the conditions was randomized. Each condition consisted of 64 trials. For the degraded condition, participants completed four repetitions of all parameter combinations: Arrow location (top vs. bottom) X central arrow direction (left-up vs. left-down vs. right-up, vs right-down) X congruency: (congruent vs. incongruent). For the standard and enhanced condition, participants completed eight repetitions of all parameter combinations to match the number of trials in the degraded condition: Arrow location (top vs. bottom) X central arrow direction (left-up vs. left-down vs. right-up, vs right-down) X congruency: (congruent vs. incongruent). Across all three conditions, the arrows were randomly presented in the top-half or bottom-half of the screen at 200 pixels (2 inches) from the center of the screen. This was done to prevent participants from rigidly fixating on the center of the screen, which might reduce the interfering effect of the flanking arrows. Each trial started with a fixation cross presented at the center of the screen. After a delay of 1000ms, the arrows are presented either in the top- or bottom-half of the screen.

Prior to each block, participants performed eight practice trials of the current condition. During practice, participants received performance feedback after each trial (either "Correct!" printed in green, "Incorrect!" printed in red, or "Too slow!" printed in red if their response latency exceeded 2000 ms). After completing the practice trials, participants completed each block with the opportunity to take a break in between. At the onset of each block, participants were told what the current condition was. No performance feedback of any kind was provided during the test blocks. The main performance measures were RTs and accuracy for each condition.

@@

Perceived unpredictability before age 13 was measured using two different scales: A scale of perceived childhood unpredictability used in previous research [@mittal_2015; @young_2018], an adaptation of the Questionnaire of Unpredictability in Childhood [QUIC; @glynn_2019], and two scales measuring change in the family and social environment. The perceived childhood unpredictability scale contained eight items measuring perceived unpredictability before age 13 (e.g, "My family life was generally inconsistent and unpredictable from day-to-day"; See [Table S3](#appendix) for an overview of all items). Participants rated each on a scale from 1 (never true) to 5 (very often true).

We included the QUIC because it captures several dimensions of environmental and household unpredictability. The QUIC distinguishes between items capturing more short-term unpredictability (i.e., on the level of seconds to minutes) and more long-term unpredictability (i.e., on the level of days to months). The scale of perceived childhood unpredictability combines items on different time scales. Some items measure unpredictability on a relatively short-term scale (e.g., item 1) while other items measure unpredictability on a long-term scale (e.g., item 3). Unpredictability on different time scales might have different effects on the development of specific attention styles.

We included all five scales of the QUIC: 1) Parental monitoring and involvement, 2) Parental predictability, 3) Parental environment, 4) Physical environment, and 3) Safety and security (See [Table S4](#appendix) for an overview of all items). The items of the Parental environment subscale deviated most from the original.

We made three general changes to the original scale as described in @glynn_2019. First, we adapted all items to refer to experiences before age 13. This was done to reduce cognitive load from having to go back-and-forth between different time scales. Second, most items were rated on a scale of 1 (never true) to 5 (very often true) instead of the original yes/no answer format. An exception was made for four items of the parental environment scale which asked for more specific experiences (e.g., "I experienced changes in my custody arrangement"; see [Table S4](#appendix)). For these items, we adopted a response scale with the options "never", "only once", "a couple times", "several times", "many times". Third, quantifiers such as "frequently", "often", and "There was a period of time when [...]" were dropped to better match the response scale. We excluded the item "My parents got divorced" because it did not fit the response format and was captured by one of the items of the perceived unpredictability scale.

We assessed household chaos before age 13 using an adaptation of the Confusion, Hubbub, and Order Scale [CHAOS; @matheny_1995]. The CHAOS consists of 15 items measuring the level of chaos in the household (e.g., "No matter how hard we tried, we always seemed to be running late"; See [Table S5](#appendix) for an overview of all items). We made two changes to the original scale as described in @matheny_1995. First, all items were converted from the present tense to the past tense, and were endorsed as applying to participants' lives before age 13. Second, all items were rated on a scale of 1 (never true) to 5 (very often true) instead of the original yes/no answer format. This change was also implemented to reduce cognitive load by keeping the answer options the same between scales.

In addition, we measured the stability of the family and social environment. On a scale of 1 (the same all the time) to 5 (constant and rapid changes), participants indicated how often the following aspects of their family and social environment changed before age 13: 1) economic status, 2) family environment, 3) childhood neighborhood environment, and 4) childhood school environment. Finally, we included several objective proxy measures of unpredictability before age 13: 1) "How often did you move", 2) "Before age 13, did any other adults besides your caregivers live in your primary residence?". If the participant answered "yes" to this question, two additional questions asked about the number of different adult women (men) that lived in their primary residence for a period of time besides their primary caregiver. 3) If the participant indicated more adults living in the household besides their primary caregiver, this was followed up by items asking how many of these individuals were romantically involved with their caregiver.

four objective measures of unpredictability before age 13. Participants provided answers to the following items in an open response format: 1) "How often did you move?"; 2) "How many adults lived in your home on average?"; 3) "Think about the caregiver who had the most romantic partners before you were age 13. How many romantic partners did this person have?".

We computed four main unpredictability variables to be able to assess how they would relate to the dependent variables as well as to each other: 1) *unpredictability~perceived~*: containing all items of the perceived unpredictability scale and the QUIC, given their close conceptual overlap (*M* = `r #TBD`, *SD* = `r #TBD`, $\alpha$ = `r #TBD` *r* = `r #TBD` between the two scales); 2) *unpredictability~CHAOS~*: containing the items of the CHAOS (*M* = `r #TBD`, *SD* = = `r #TBD`, $\alpha$ = `r #TBD`); 3) *unpredictability~subjective~*: containing all items of the perceived unpredictability scale, the QUIC and the CHAOS (*M* = `r #TBD`, *SD* = `r #TBD`, $\alpha$ = `r #TBD`); 4) *unpredictability~objective*: containing the items assessing the stability of the family- and social environment, as well as the three objective measures (*M* = `r #TBD`, *SD* = `r #TBD`, $\alpha$ = `r #TBD`).

We computed several unweighted unpredictability variables to be able to assess how they would relate to the dependent variables as well as to each other: 1) overall perceived unpredictability variable containing all eight items of the perceived unpredictability scale (*M* = `r #TBD `, *SD* = = `r #TBD `, $\alpha$ = `r #TBD `); 2) an overall QUIC variable by averaging all 37 items (*M* = `r #TBD `, *SD* = = `r #TBD `, $\alpha$ = `r #TBD `); 3) five separate QUIC subscales by averaging their sub-items (Parental monitoring and involvement: *M* = `r #TBD `, *SD* = `r #TBD `, $\alpha$ = `r #TBD `; Parental predictability: *M* = `r #TBD `, *SD* = `r #TBD `, $\alpha$ = `r #TBD `; Parental environment: *M* = `r #TBD `, *SD* = `r #TBD `, $\alpha$ = `r #TBD `; Physical environment: *M* = `r #TBD `, *SD* = `r #TBD `, $\alpha$ = `r #TBD `; Safety and security: *M* = `r #TBD `, *SD* = `r #TBD `, $\alpha$ = `r #TBD `). See also [below](#efa) for details on a factor analysis based on the individual items across scales.

@@

Participants' perceived level of resource scarcity before age 13 was measured using seven items (e.g., "Your family had enough money to afford the kind of home you all needed"; See [Table S6](#appendix) for an overview of all items). Participants rated each item on a scale from 1 (never true) to 5 (very often true). Scores for the first six items were reverse coded so that higher scores indicated more perceived resource scarcity. The items were averaged together to create an unweighted composite scale (*M* = `r  #TBD `, *SD* = `r #TBD `, $\alpha$ = `r #TBD `).

In addition, we measured several indicators of objective SES before age 13. First, participants separately indicated the highest education of their two primary caregivers (if applicable) on an 8-point scale: 'some high school', 'GED', 'high school diploma', 'some college but no college degree', associate's  degree', 'bachelor's degree', 'master's  degree', or 'doctoral or lab degree'. The caregivers' education level were averaged to create an overall unweighted parental education composite (*M* = `r  #TBD `, *SD* = `r  #TBD `). Participants also indicated their family's household income before age 13 on a 6-point scale: 'less than \$ 25k/year', '\$25k - \$49k/year, '\$50 - \$74k/year', '\$75 - \$99k/year', '\$100 - \$149k/year', 'more than \$150k/year'. Scores were reverse coded so that higher scores indicated higher levels of poverty.

We created a composite score of poverty exposure before age 13 by averaging together the standardized scores of perceived level of resource scarcity, overall caregiver education, and household income.

@@

We assessed impulsivity with the Motor Impulsivity subscale of the Barrett Impulsivity Scale (BIS; short form; [@patton_1995; @spinella_2007]). The Motor Impulsivity subscale of the BIS consists of five items (e.g., "I do things without thinking"; See [Table S7](#appendix) for an overview of all items). We did not include the Non-planning subscale because it overlapped substantially with the Future Orientation Scale described below. In addition, we did not include the Attention impulsivity subscale because it included items which we deemed to be mostly irrelevant for our target population (e.g., "I 'squirm' at plays or lectures"). We changed the original 4-point rating scale (rarely/never to almost always) to a 5-point rating scale ranging from 1 (never true) to 5 (very often true). An overall impulsivity variable was computed by averaging the five unweighted items (*M* = `r #TBD `, *SD* = `r #TBD `, $\alpha$ = `r #TBD `).

@@

We assessed future orientation with an adapted version of the Future Orientation Scale [FOS; @steinberg_2009]. The original scale consists of 15 sets of opposing items separated by "BUT" (e.g., "Some people like to plan things out one step at a time BUT other people like to jump right into things without planning them out beforehand"). Participants first choose the item that best matches their general preference, and then indicate whether the statement is "really true" or "somewhat true". We adapted this format in a couple of ways. First, we converted the two statements per item to a single statement by picking the statements in the original right-hand column. Second, we adapted the 15 statements from a third-person to a first-person format. These changes were made in an attempt to reduce the cognitive load of the items. We worried that people with less formal education or who were sitting in a noisier environment would struggle with the length of the original items.

In addition, item 8 of the original scale ("[...] other people would rather spend their money right away on something fun than save it for a rainy day") was changed to "I'd rather spend money right away than save it for a rainy day" (i.e., dropping the phrase "on something fun") to make it more general with regard to the thing that money is spent on. For people from adversity, spending money right now instead of saving it for the future might often be born out of necessity (e.g., having just enough money for food and shelter; being in debt) instead of a failure to delay gratification. See [Table S8](#appendix) for an overview of all adapted items). Finally, the rating scale was adapted from the original 4-point scale (ranging from really true for the left-hand statement to really true for the right-hand statement) to a 5-point scale ranging from 1 (never true) to 5 (very often true).

An overall future orientation variable was computed by averaging the 15 unweighted items (*M* = `r #TBD `, *SD* = `r #TBD `, $\alpha$ = `r #TBD `). In addition, we calculated subscales for "Planning ahead" (items 1, 6, 7, 12 and 13), "Time perspective" (items 2, 5, 8, 11 and 14) and "Anticipation of future consequences" (items 3, 4, 9, 10 and 15).

@@

We assessed depressive symptoms during the past week using the Center for Epidemiologic Studies Depression Scale [CESD; @radloff_1977]. The scale consists of 20 items (e.g., "I do things without thinking"; See [Table S9](#appendix) for an overview of all items). Participants rate each item on a scale of 1 (rarely or none of the time (less than 1 day)) to 4 (most or all of the time (5-7 days)). An overall depression variable was computed by averaging the 20 unweighted items (*M* = `r #TBD `, *SD* = `r #TBD `, $\alpha$ = `r #TBD `).

@@

The experiment was completed on the participants' own laptop or desktop computer and consisted of five parts (in fixed order): consent, Flanker Task, questionnaire battery, brief demographics form, and final checks including the opportunity to give feedback on the experiment. Participants were allowed to refrain from answering any of the questionnaire items, but were prompted with a warning once when moving to the next page if one of the items was not answered (which they could ignore).

After providing consent, participants started with the Flanker Task. They were asked to complete the Flanker Task in a quiet room in the house where they would be least likely to be distracted by other people or outside noises. The order of the tasks was counterbalanced between subjects. At the onset of the first task, the experiment went into full-screen mode to limit distractions from other programs or browser tabs. The size of the task stimuli was controlled between subjects using the resize plugin in JsPsych [@deLeeuw_2015]. Participants were asked to hold a creditcard (or similarly sized card) up against the screen and to increase the size of a blue rectangle on the screen (initial size 300 pixels) until it matched the size of the creditcard. The stimulus display for each task was resized so that 100 pixels corresponded to 1 inch for all participants. After successfully resizing the screen, participants completed all three tasks. During the task, the cursor was hidden from the screen to minimize distractions. Code for the three tasks can be found on [Github](https://github.com/StefanVermeent/attention_project/tree/main/tasks).

After completing the Flanker task, participants completed the questionnaire battery in the following fixed order: 1) Current state (state anxiety and separate questions relating to specific states); 2) Childhood adversity (perceived unpredictability, perceived socio-economic status, exposure to violence and physical fights, household chaos); 3) Temporal orientation (impulsivity, future orientation); 4) Depressive symptoms.

Finally, the demographics questions asked about the participant's age, weight, height, physical activity, sex at birth, gender, ethnicity, social class (current and during childhood), education level (one's own as well as caregivers' education level), occupation, and household income (current and during childhood). At the end of the experiment, we asked participants if they ever got up or were interrupted during the study, and how noisy their environment was during the attention tasks.

The full experiment was expected to take approximately 30 minutes. Participants were paid £3.75 upon completing the full experiment.

@@

The DDM analysis of the Flanker task was done using an adaptation of the standard DDM that was specifically developed to account for Flanker data: The Shrinking Spotlight model [SSP; @grange_2016; @white_2018a; @white_2018b; @white_2011]. The standard DDM model assumes a stable drift rate over time within trials (i.e., the rate of information accumulation does not change over time). However, on the Flanker Task, reaction time patterns indicate that Flankers exert their strongest effect on performance early on the trial, which then decreases over time. This effectively means that the drift rate increases over time, thus violating the basic assumption of the standard DDM [@white_2011]. The SSP explains this pattern by assuming that attention on the Flanker task works like a spotlight that starts relatively broad, and then gradually narrows down to the central arrow over time. Each arrow provides perceptual input *p*. If the flanking arrows are congruent, all flankers take a positive value for *p*; if the flanking arrows are incongruent, they take a negative value for *p* while the central arrow takes a positive value. The drift rate (*v*) at each time point is the function of the strength of perceptual input *p* multiplied by the amount of attention focused on each arrow.

Attention is operationalized by two additional parameters: the initial width of the attentional spotlight (*SD~a~*) and the shrinking rate of the attentional spotlight (*rd*). The attentional spotlight is assumed to be normally distributed over the arrows, with relatively more attention paid to the arrows at the center and relatively less attention paid to the peripheral arrows. Over time, this normal distribution (*SD~a~*) narrows down to the central arrow at the rate specified by *rd*, thereby gradually decreasing the interfering noise of the flankers and increasing the signal obtained from the target.

Model fit was done using the *flankr* package in R [@grange_2016]. The SSP model was fit to the data of each individual participant in two steps. First, we looped over 50 different sets of starting parameters with a variance of 20 to find the best starting values for each participant. This was done to prevent accidentally converging on a suboptimal solution [@grange_2016]. On this first fitting step, 1000 trials were simulated on each iteration of the fit routine. Second, once the optimal set of starting values was found, we fitted the final model using the best starting values found in step 1 and simulating 50,000 trials on each iteration.

The model fit for all participants was assessed through QQ-plots comparing the empirical data with the model predictions following @grange_2020. Using each participant's best-fitting parameters (based on the final step above), we simulated RTs and accuracy for 50,000 trials of each congruency condition and each Flanker version. We calculated the total proportion accuracy and the 25th, 50th, and 75th quantile of the response time distribution (correct trials only) for both the simulated data and the empirical data. The outcomes for all participants were plotted against each other. The closer the points are to the diagonal line, the better the models fit the participants' data.

For the analyses, we extracted the following parameters: 1) perceptual input (*p*); 2) boundary separation (*A*); 3) initial attention width (*sd~a~*); 4) Attention shrinking rate (*rd*); Interference effect (*sd~a~* / *rd*); 5) non-decision time (*t0*).

@@

Online experiments come at the trade-off of having less experimental control over the way in which participants complete the experiment and over the environment in which they do so. It is largely unclear which factors affect performance and how strong these effects are. We used multiverse analysis for all main analysis to assess the robustness of the results against various environmental factors and situations during the experiment. We identified seven arbitrary analytic decisions, including or excluding 1) participants who had a recaptcha score below 0.5 (possibly indicating bots); 2) participants who did not rescale their screen at the start of the experiment (see the [procedure section](#procedure)); 3) participants who did not enter fullscreen mode prior to starting the Flanker Task; 4) participants who exited fullscreen mode at any point during the Flanker Task; 5) participants who indicated high levels of noise in their environment; indicated extreme interruptions during the experiment; 7) trials with RTs > 3.2 SD (within subjects).

For each analysis, we report the median $\beta$, 95% confidence intervals, proportion of *p*-values < .05 across all analytic decisions. For the primary analyses, we used a bootstrapping technique to compute overall *p*-values to assess whether the obtained median $\beta$ is significantly larger than zero [@simonsohn_2020].

### Primary analyses {#primary_plan}

>>>>

Generally speaking, the results of the first two studies were not in line with our initial predictions about a present-oriented attention style.
Although we expected such a style to lead to faster speed of perceptual processing on attention tasks, we generally found the opposite pattern.
One reason for this might be that a present-oriented attention style leads people from adversity to process information more holistically, which would lower the depth of perceptual processing of any individual stimulus.
Another related reason might be that an external, present-oriented attention style extends beyond the computer screen to all of their immediate environment.
Both of these are potential mechanistic explanations for a lowered quality of perceptual processing that would be in line with a specialized attention style.
In study 2, we aimed to replicate the broader findings from study 1 (lowered perceptual input on the Flanker Task but no differences in interference) and to investigate if and how the lower quality of perceptual processing was associated with a more holistic style of attention.

@@

1. Investigate how unpredictability and violence exposure are associated with perceptual input and interference on the Flanker Task by pooling the data across all three studies.

2. Investigate how unpredictability and violence exposure are associated with a preference for more holistic vs. detailed processing of information.

@@

Participants will be 400 US-based individuals between the ages of 18 and 30. Recruitment will be identical to study 1. 
We conducted a power analysis for the planned linear regressions through simulation. 
We determined the required sample size to achieve power of > .90 for a standardized main effect of 0.05.
This main effect is slightly lower than the effect of 0.06 that we found for perceptual input in the pooled analyses of study 1.
Simulations revealed that we would require around 450 participants to achieve a power of .93.
We decided to recruit 500 participants with the expectation that we will have to exclude around 50 participants.

@@

Participants will be excluded if they 1) do not complete the full study or both cognitive tasks, 2) miss both attention check items or if they have suspicious response patterns (e.g., consistently endorsing high response options even when some items were reverse coded), 3) have a screen height < 700 pixels or a screen height that is bigger than the screen width to ensure the tasks were completed on a laptop or desktop pc, 4) exit full-screen mode and/or engaged with other broswer tabs *during* the cognitive tasks (but not while reading instructions or while taking breaks in between blocks), 5) do not perform above chance on either the Flanker Task or the Global-Local Task (based on a binomial distribution using a probability of 97.5%).

In addition, we will screen reaction times on both cognitive tasks. 
We will remove trials with response times < 250 or > 3500 ms.
Participants with more than 10 removed trials will be excluded from the analyses.

@@

@@

The Flanker Task was identical to the standard version used in study 1. 
It was programmed in JsPsych 7.3.2 [@deleeuw_2015]. 
Participants first completed eight practice trials and next completed 64 test trials in a single block.

## Global-Local Task {#globallocal}

@@

We distinguish between subjective and objective measures of environmental unpredictability.
We distinguish between subjective and objective measures of environmental unpredictability.
Subjective unpredictability will be measured as an unweighted average of three scales: 1) A scale of perceived childhood unpredictability used in previous research (see [Table S3]) [@mittal_2015; @young_2018]; 2) An adapted version of the Confusion, Hubbub and Order Scale [CHAOS; @matheny_1995] (see [Table S5](#appendix)); 3) An adapted version of the Questionnaire of Unpredictability in Childhood [QUIC; @glynn_2019] (see [Table S4](#appendix)).
Objective unpredictability will be measured as an unweighted average of four items relating to experiences prior to age 13: 1) number of residential changes, 2) number of male and female romantic partners of a caregiver besides the other primary caregiver (measured separately; values > 6 were given a value 6 to prevent extreme outliers), 4) self-reported stability of the family and social environment across four dimensions: economic status, family environment, childhood neighborhood environment, and childhood school environment (rated between 1 (the same all the time) and 5 (constant and rapid changes)).
The main analyses will be based on a full composite that is an unweighted standardized average of subjective and objective unpredictability.

@@

We will create a poverty composite score that is the average of the following measures (standardized) focusing on the family situation prior to age 13: 1) perceived level of resource scarcity (See [Table S6](#appendix)); 2) the unweighted average of the highest achieved education of both primary caregivers; 3) household income (6-point scale: 'less than \$ 25k/year', '\$25k - \$49k/year, '\$50 - \$74k/year', '\$75 - \$99k/year', '\$100 - \$149k/year', 'more than \$150k/year'. Scores were reverse coded so that higher scores indicated higher levels of poverty.).

### Attentional style

We will measure attentional style using the Attentional Style Questionnaire [ASQ; @calster_2018].
The ASQ measures self-reported attentional style, with seven items

@@

We will measure impulsivity with the Motor Impulsivity subscale of the Barrett Impulsivity Scale (BIS; short form; [@patton_1995; @spinella_2007]) (See [Table S7](#appendix)) 
An overall impulsivity variable will be computed by averaging the five unweighted items.

@@

We will measure future orientation with an adapted version of the Future Orientation Scale [FOS; @steinberg_2009] (See [Table S8](#appendix).
An overall future orientation variable will be computed by averaging the 15 unweighted items.

@@

We will measure depressive symptoms during the past week using the Center for Epidemiologic Studies Depression Scale [CESD; @radloff_1977] (See [Table S9](#appendix). 
An overall depression variable will be computed by averaging the 20 unweighted items.

@@

The experiment will be completed on the participants' own laptop or desktop computer and will of five parts (in fixed order): consent, cognitive tasks, questionnaire battery, brief demographics form, and final checks including the opportunity to give feedback on the experiment. Participants are allowed to refrain from answering any of the questionnaire items, but are prompted with a warning once when moving to the next page if one of the items is not answered (which they can ignore).

After providing consent, participants start either with the Flanker Task or Global-Local Task (counterbalanced). They are asked to complete the tasks in a quiet room in the house where they would be least likely to be distracted by other people or outside noises. 
At the onset of the first task, the experiment switches to full-screen mode to limit distractions from other programs or browser tabs.
The size of the task stimuli is controlled between subjects using the resize plugin in JsPsych [@deLeeuw_2015]. 
After successfully resizing the screen, participants complete all three tasks. 
Code for the three tasks can be found at [https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/3_study2/materials](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/3_study2/materials).

After completing the cognitive tasks, participants complete the questionnaire battery in the following fixed order: 1) Current state (state anxiety and separate questions relating to specific states); 2) Childhood adversity (perceived unpredictability, perceived socio-economic status, exposure to violence and physical fights, household chaos); 3) Temporal orientation (impulsivity, future orientation); 4) Depressive symptoms.

Finally, the demographics questions asks about the participant's age, weight, height, physical activity, sex at birth, gender, ethnicity, social class (current and during childhood), education level (one's own as well as caregivers' education level), occupation, and household income (current and during childhood). 
At the end of the experiment, we ask participants if they ever got up or were interrupted during the study, and how noisy their environment was during the attention tasks.

The full experiment is expected to take approximately 30 minutes. Participants are paid £3.75 upon completing the full experiment.

@@

For the Flanker Task, we will use the Shrinking Spotlight Model [SSP; @grange_2016; @white_2018a; @white_2018b; @white_2011] to fit the data.
The SSP model will be fit to the data of each individual participant in two steps. 
First, we loop over 50 different sets of starting parameters with a variance of 20 to find the best starting values for each participant.  
On this first fitting step, 1000 trials will be simulated on each iteration of the fit routine. 
Second, after finding the optimal set of starting values, we fit the final model using the best starting values found in step 1 and simulate 50,000 trials on each iteration. 
The main parameters of interest are perceptual input (*p*) and interference (initial attention width / attention shrinking rate).

For the Global-Local Task, we will use an hierarchical Bayesian DDM to fit the data using the runjags package [@denwood_2016].
Drift rate and non-decision time will be freely estimated per task condition (global vs. local), whereas boundary separation wil be fixed to be equal across conditions.
The starting point will be fixed to 0.5.
Each model will be fit with three Markov Chain Monte Carlo (MCMC) chains. Each chain will contain 2,000 burn-in samples and 10,000 additional samples. 
Of these samples, every 10th sample will be retained. 
Posterior samples of all three chains will be combined, resulting in a posterior sample of 3,000 samples. 
The main parameter of interest is the drift rate, more specifically, the difference between drift rates (drift rate~global~ - drift rate~local~).

### Primary analyses

To address the [first primary aim](#primary), we will fit a linear regression model to the pooled data of all three studies (i.e., the pilot study, study 1, and study 2).
We will fit different models for environmental unpredictability and violence exposure as predictors (using the respective composite scores) and for perceptual input and interference as dependent variables.
Each model will additionally contain the following covariates: study number (...), whether participants rescaled their screen at the start of the cognitive tasks (yes or no, sum-coded), whether participants experienced interruptions during the tasks (yes or no, sum-coded), and whether people exited full-screen mode during the experiment (yes or no, sum-coded).
Note that the latter three covariates were included as arbitrary decisions in the multiverse analyses for the pilot study and study 1.
However, as we found that they substantially altered the results, we decided to reclassify them as non-arbitrary decisions and include them as covariates instead.
To address the [second primary aim](#primary), we will fit the same model as described above for the first primary aim, but with different the difference score of the Global-Local drift rate as the dependent variable.

@@

We will conduct a multiverse analysis for all primary analyses.
We identified seven arbitrary analytic decisions, including or excluding 1) participants who had a recaptcha score below 0.5 (possibly indicating bots); 2) participants who did not enter fullscreen mode prior to starting the cognitive tasks; 3) participants who indicated high levels of noise in their environment.
For each analysis, we report the median $\beta$, 95% confidence intervals, proportion of *p*-values < .05 across all analytic decisions. For the primary analyses, we used a bootstrapping technique to compute overall *p*-values to assess whether the obtained median $\beta$ is significantly larger than zero [@simonsohn_2020].
